---
title: 'Stat 579 - Homework #4 - Section A'
author: "Your Name"
date: "9/26/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Behavioral Risk Factor Surveillance System - Again

We are again looking at the data from the
Behavioral Risk Factor Surveillance System (BRFSS).

A subset of the data concentrating on Iowa with records for 2016 is given at
```{r}
url <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/brfss-iowa-2018.csv"
```

The following code reads the data into your R session:
```{r, cache = TRUE, warning = FALSE}
iowa <- read.csv(url)
library(tidyverse)
```

A codebook describing the survey and a listing of all variables is available at [https://www.cdc.gov/brfss/annual_data/2017/pdf/codebook17_llcp-v2-508.pdf]().

For each of the questions, show the code necessary to get to the answer. 
Make sure to also write the answer to the question in a sentence.

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. Load the dataset into your session and store it in the object `iowa`.

As already seen in the last homework, the coding scheme used for HEIGHT3 is:

| HEIGHT3 value  |  Interpretation    |
|:- |:- |
| 200 - 711 | Height (ft/inches), i.e. 410 is 4 feet, 10 inches	|
| 7777 |	Don't know/Not sure	|
| 9000 - 9998 |	Height (meters/centimeters), where the first 9 indicates that the measurement was metric, 9165 is 1 meter 65 cm |
| 9999	| Refused |
| BLANK |	Not asked or Missing |

The overall goal is to introduce a new variable `height` that corresponds to reported height in centimeters [cm].

3. Sketch out a way to derive a new variable `height` that represents the values in `HEIGHT3` in centimeters and properly encodes any unavailable information as missing values (`NA`).
Implement your transformation. For your convenience: 1 ft equals 30.48 cm, 1 inch equals 2.54 cm. 1 m equals 100 cm.
```{r}
iowa <- iowa %>%
  mutate(feet = HEIGHT3 %/% 100)
for (i in 1:length(iowa$feet)) {
  if (iowa$feet[i] >= 77 | is.na(iowa$feet[i])) {
    iowa$feet[i] <- NA
  }
}
sum(is.na(iowa$feet))
iowa <- iowa %>%
  mutate(inch = HEIGHT3 %% 100)
for (i in 1:length(iowa$feet)) {
  if (is.na(iowa$feet[i])) {
    iowa$inch[i] <- NA
  }
}
sum(is.na(iowa$inch))

iowa <- iowa %>%
  mutate(height = feet*30.48 + inch*2.54)

iowa <- iowa %>% 
  mutate(metric = ifelse(between(HEIGHT3, 9000, 9998), HEIGHT3 %% 1000, NA))

for (i in 1:dim(iowa)[1]) {
  if (!is.na(iowa$metric[i])) {
    iowa$height[i] <- iowa$metric[i]
  }
}


```

4. How  many missing values does the variable `height` have? 

```{r}
sum(is.na(iowa$height))
```
### There are 252 missing values

5. What is the average height among respondents of the survey? among men? among women? Does the result make sense? Comment.

```{r}
mean(iowa$height, na.rm=T)
mean(iowa$height[iowa$SEX1==1], na.rm=T)
mean(iowa$height[iowa$SEX1==2], na.rm=T)
```
### The average height among respondents is 171.228cm. Among men is 178.821cm. And among women is 164.275cm.

6. Plot a histogram of `height`, facet by gender such that the histograms are on top of each other. Choose an appropriate bin width and comment on the resulting plot. 
```{r}
iowa %>% 
  ggplot(aes(x=height,fill = factor(SEX1)) ) + 
  geom_histogram() #+
  facet_wrap(~SEX1)
```
7. The variable `WTKG3` has measurements of weight in 10s of grams (i.e. `WTKG3` divided by 100 gives the weight in kilogram). 


8. Use `ggplot2` to draw a scatterplot of weight (`WTKG3`) by height (new `height` variable). Facet by gender and comment.
```{r}
iowa %>% 
  ggplot(aes(x = height, y = WTKG3)) +
  geom_point() +
  facet_wrap(~SEX1)
```
Due date: the homework is due a week from now on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
